\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
eBoxed the Gordan Ramsey chef kiss

After using the defining Jacobi relations to relate every graph to the
basis, any remaining Jacobi relations will be referred to as boundary
relations.  In other words, boundary relations (indirectly) relate
basis elements to themselves.  For example, the crossed box numerator
is related to itself via the following Jacobi relation \be % [[19]]
\label{eq:TwoLoopJacobiTwo}
\JacobiTwoCrossedBoxOne + \JacobiTwoCrossedBoxTwo +  \JacobiTwoCrossedBoxThree =0,
\end{equation}
where the shared propagator is shown in red as before.  Each of the
crossed box numerators can be written in terms of the double box and
penta-triangle so this Jacobi relation places constraints on the basis
numerator \ansatze{}.  While defining Jacobi relations simplify how
many numerators require \ansatze{}, boundary relations place
restrictions on those \ansatze{}.  In principle there could be Jacobi
relations involving tadpoles and BEL graphs, which would be subtle to
handle.  One option is to force the numerators of these graphs to
zero.  However, since these graphs will not contribute to physical
unitarity cuts, we take a different approach.  These graphs are
allowed to have non-zero numerators but any Jacobi relations involving
them are simply not imposed or solved.  See Refs.~\cite{Bern:2012uf,
  Edison:2022jln} and \cref{BELref} for discussions of BELs in the context of
color-kinematics duality.

@@

In fact, we can make the failure extremely precise.  Start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}:
the double-box, crossed-box, and penta-triangle.  For each of the
three diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop).  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2761 free parameters on the double-box, 2576 free parameters on the
crossed-box, and 5040 free paramters on the penta-triangle.  Then
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.

@@

Here we apply the same conventions of Ref.~\cite{jjmcTASI2014}, which we quote now. The component definition of our spinor bracket that are consistent with the conventions above,
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\end{align}
where the $a_i$ are component values of the four-vector, $k^\mu_a = (a_0,a_1,a_2,a_3)$. For massless momenta, $k_a$ and $k_b$, one can apply the on-shell condition, $k_a^2=k_b^2=0$, and show that:
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]\,.
\end{equation}
Four-dimensional polarization dot products with fixed helicity states can be mapped as follows:
\begin{equation}\label{eq:4DPols}
\begin{aligned}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}\,,
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}\,,
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} \,,
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 \,,
\end{aligned}
\end{equation}
Note that given the above definition, the spinor helicity variables carry the same mass dimension for both the angle and square brackets. In terms of light-cone coordinates that we employed above in \cref{sec:CKLagrangians}, the angles and squares can be rewritten as follows:
\begin{align}
\langle ab \rangle &= \frac{a_w b_u-a_u b_w}{\sqrt{a_ub_u}}\,,
\\
[ab] &= \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{\sqrt{a_ub_u}}\,,
\end{align}
The conventions we use in the text where the holomorphic spinor, $X(p,k)$, carries extra mass dimension essentially amounts to shifting the bracket by the following overall factor:
\begin{align}
\langle ab \rangle &\rightarrow a_w b_u-a_u b_w \equiv X(a,b)\,,
\\
[ab] &\rightarrow \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{a_ub_u}\equiv Q(a,b)\,,
\end{align}
With this, we still obtain the same completeness relation for the SDYM spinors:
\begin{equation}
X(a,b)Q(b,a) = \frac{1}{2}s_{ab}
\end{equation}
\section{Regulating BEL integrals}\label{BELreg}
At one-loop, weight counting tells us that the $n$-gon master numerator must have $n$ on-shell $Z$-particles. Unitarity requires that there are three distinct contributions from $Y\!Z$ theory -- the first from an off-shell $Y$-loop particle and then two more from different orientations of a $\bar{Z}Z$-loop.  Thus, $YZ$ theory gives us the following one-loop $n$-gon numerator:
\begin{equation}
N^{n\text{-gon},YZ}_{(12...n)} = (T^{1}T^{2}\cdots T^{n})+ (F^1F^2\cdots F^n)
\end{equation}
where $(\,\cdots)$ indicates an internal contraction over the $YY$ and $\bar{Z}Z$ loops. Plugging in the Feynman rule of \cref{eq:FeynmanRuleYYZ} and \cref{eq:FeynmanRuleXZZ}, we can readily obtain expressions in terms of the internal loop factors $l_i$ and the external momenta $k_i$ (we define the $l_i$ loop momentum as that flowing into $k_i$ and out of $k_{i-1}$):
\begin{equation}
N_{(12...n)}^{\text{NLSM}}=(T^{1}T^{2}\cdots T^{n}) = (l_1 k_1)(l_2 k_2) \cdots (l_n k_n)
\end{equation}
and similarly so for the internal vector contribution:
\begin{equation}
 (F^1F^2\cdots F^n) = D (l_1 k_1)(l_2 k_2) \cdots (l_n k_n) + \mathcal{O}(D^0)
\end{equation}
The dimension dependent factor essentially counts that number of internal vector states. While this $n$-gon is manifestly color-dual, it does not produce the right cuts for NLSM. However, the scalar contribution, that comes dressed with an overall factor $D$ \textit{does} manifest the duality globally, and satisfies all the desired pion cuts. In order for the Feynman rules for $Y\!Z$ theory compute one-loop color-dual numerators consistent with NLSM cuts, we must add some additional states to cancel off he spurious poles, while preserving color-kinematics duality. We leave this as a direction of future work.

>>>>

After using the defining Jacobi relations to relate every graph to the
basis, any remaining Jacobi relations will be referred to as boundary
relations.  In other words, boundary relations (indirectly) relate
basis elements to themselves.  For example, the crossed box numerator
is related to itself via the following Jacobi relation \be % [[19]]
\label{eq:TwoLoopJacobiTwo}
\JacobiTwoCrossedBoxOne + \JacobiTwoCrossedBoxTwo +  \JacobiTwoCrossedBoxThree =0,
\end{equation}
where the shared propagator is shown in red as before.  Each of the
crossed box numerators can be written in terms of the double box and
penta-triangle so this Jacobi relation places constraints on the basis
numerator \ansatze{}.  While defining Jacobi relations simplify how
many numerators require \ansatze{}, boundary relations place
restrictions on those \ansatze{}.  In principle there could be Jacobi
relations involving tadpoles and BEL graphs, which would be subtle to
handle.  One option is to force the numerators of these graphs to
zero.  However, since these graphs will not contribute to physical
unitarity cuts, we take a different approach.  These graphs are
allowed to have non-zero numerators but any Jacobi relations involving
them are simply not imposed or solved.  See Refs.~\cite{Bern:2012uf,
  Edison:2022jln} and \cref{BELreg} for discussions of BELs in the context of
color-kinematics duality.

@@

In fact, we can make the failure extremely precise.  Start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}:
the double-box, crossed-box, and penta-triangle.  For each of the
three diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop).  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2761 free parameters on the double-box, 2576 free parameters on the
crossed-box, and 5040 free paramters on the penta-triangle.  Then
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{eBox}
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
\end{eBox}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.

@@

Here we apply the same conventions of Ref.~\cite{jjmcTASI2014}, which we quote now. The component definition of our spinor bracket that are consistent with the conventions above,
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\end{align}
where the $a_i$ are component values of the four-vector, $k^\mu_a = (a_0,a_1,a_2,a_3)$. For massless momenta, $k_a$ and $k_b$, one can apply the on-shell condition, $k_a^2=k_b^2=0$, and show that:
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]\,.
\end{equation}
Four-dimensional polarization dot products with fixed helicity states can be mapped as follows:
\begin{equation}\label{eq:4DPols}
\begin{aligned}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}\,,
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}\,,
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} \,,
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 \,,
\end{aligned}
\end{equation}
Note that given the above definition, the spinor helicity variables carry the same mass dimension for both the angle and square brackets. In terms of light-cone coordinates that we employed above in \cref{sec:CKLagrangians}, the angles and squares can be rewritten as follows:
\begin{align}
\langle ab \rangle &= \frac{a_w b_u-a_u b_w}{\sqrt{a_ub_u}}\,,
\\
[ab] &= \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{\sqrt{a_ub_u}}\,,
\end{align}
The conventions we use in the text where the holomorphic spinor, $X(p,k)$, carries extra mass dimension essentially amounts to shifting the bracket by the following overall factor:
\begin{align}
\langle ab \rangle &\rightarrow a_w b_u-a_u b_w \equiv X(a,b)\,,
\\
[ab] &\rightarrow \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{a_ub_u}\equiv Q(a,b)\,,
\end{align}
With this, we still obtain the same completeness relation for the SDYM spinors:
\begin{equation}
X(a,b)Q(b,a) = \frac{1}{2}s_{ab}
\end{equation}
\section{Regulating BEL integrals}\label{BELreg}
At one-loop, weight counting tells us that the $n$-gon master numerator must have $n$ on-shell $Z$-particles. Unitarity requires that there are three distinct contributions from $Y\!Z$ theory -- the first from an off-shell $Y$-loop particle and then two more from different orientations of a $\bar{Z}Z$-loop.  Thus, $YZ$ theory gives us the following one-loop $n$-gon numerator:
\begin{equation}
N^{n\text{-gon},YZ}_{(12...n)} = \langle T^{1}T^{2}\cdots T^{n}\rangle+ \rangle F^1F^2\cdots F^n\rangle
\end{equation}
where $(\,\cdots)$ indicates an internal contraction over the $YY$ and $\bar{Z}Z$ loops. Plugging in the Feynman rule of \cref{eq:FeynmanRuleYYZ} and \cref{eq:FeynmanRuleXZZ}, we can readily obtain expressions in terms of the internal loop factors $l_i$ and the external momenta $k_i$ (we define the $l_i$ loop momentum as that flowing into $k_i$ and out of $k_{i-1}$):
\begin{equation}
N_{(12...n)}^{\text{NLSM}}=(T^{1}T^{2}\cdots T^{n}) = (l_1 k_1)(l_2 k_2) \cdots (l_n k_n)
\end{equation}
and similarly so for the internal vector contribution:
\begin{equation}
 (F^1F^2\cdots F^n) = D (l_1 k_1)(l_2 k_2) \cdots (l_n k_n) + \mathcal{O}(D^0)
\end{equation}
The dimension dependent factor essentially counts that number of internal vector states. While this $n$-gon is manifestly color-dual, it does not produce the right cuts for NLSM. However, the scalar contribution, that comes dressed with an overall factor $D$ \textit{does} manifest the duality globally, and satisfies all the desired pion cuts. In order for the Feynman rules for $Y\!Z$ theory compute one-loop color-dual numerators consistent with NLSM cuts, we must add some additional states to cancel off he spurious poles, while preserving color-kinematics duality. We leave this as a direction of future work.
